# (PART) R Package {-} 

# Install {#install}



Here, we have provided several ways to install UCSCXenaShiny R package. You can see the detailed instruction and troubleshooting information from the README file of our [Github project](https://github.com/openbiox/UCSCXenaShiny).  



## CRAN 

- https://cran.r-project.org/package=UCSCXenaShiny

```{r, eval=FALSE}
# Online
install.packages("UCSCXenaShiny")
## or specific repos
install.packages("UCSCXenaShiny", repos = c("https://openbiox.r-universe.dev", "https://cran.r-project.org"))

# Local
install.packages("UCSCXenaShiny_2.0.0.tar.gz", repos = NULL, type="source")
```

- Once the package has been installed successfully, your can load it and check its version:

```{r}
library(UCSCXenaShiny)
packageVersion("UCSCXenaShiny")
```


## Github

```{r, eval=FALSE}
# Online
devtools::install_github("openbiox/UCSCXenaShiny")

# Local
devtools::install_local(path = "UCSCXenaShiny-master.zip")
```


## Conda
```{bash, eval=FALSE}
conda install -c conda-forge r-ucscxenashiny
```

- list all of the versions of r-ucscxenashiny available on your platform:

```{bash, eval=FALSE}
conda search r-ucscxenashiny --channel conda-forge
```

## Docker
```{bash, eval=FALSE}
# install image
docker pull shixiangwang/ucscxenashiny

# run image
docker run -d --name xenashiny -p 3838:3838 shixiangwang/ucscxenashiny
```